Priors

### 2023-06-30 reports/priors.Rmd
comparison to CTIS priors

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near $20%$.", fig.width=9}

@@

```{r}
#
# Logistic transformation of the Beta CDF.
#
f.beta <- function(alpha, beta, x, lower=0, upper=1) {
  p <- pbeta(x, alpha, beta)
 return(p)
}
#
# Sums of squares.
#
delta <- function(fit, actual) sum((fit-actual)^2)
#
# The objective function handles the transformed parameters `theta` and
# uses `f.beta` and `delta` to fit the values and measure their discrepancies.
#
objective <- function(shapes, x, prob, ...) {
  fit <- pbeta(x, shapes[1], shapes[2], ...)
  # return sum of squares
  return(sum((fit, prob)^2))
}

# Solve two problems.

par(mfrow=c(1,2))
alpha <- 15; beta <- 22 # The true parameters
for (x in list(c(1e-3, 2e-3), c(1/3, 2/3))) {
  x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
  start <- (c(1e1, 1e1))            # A good guess is useful here
  sol <- nlm(objective, start, x=x, prob=x.p, lower=0, upper=1,
             typsize=c(1,1), fscale=1e-12, gradtol=1e-12)
  parms <- (sol$estimate)           # Estimates of alpha and beta
  #
  # Display the actual and estimated values.
  #
  print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
  #
  # Plot the true and estimated CDFs.
  #      
  curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
  curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
  points(x, pbeta(x, alpha, beta))
}

for (x in list(c(.1200,.2188,.3683))) {
  # x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
  start <- (c(1e1, 1e1))            # A good guess is useful here
  sol <- nlm(objective, start, x=x, prob=c(.025,.5,.975), lower=0, upper=1,
             typsize=c(1,1), fscale=1e-12, gradtol=1e-13)
  parms <- (sol$estimate)           # Estimates of alpha and beta
  #
  # Display the actual and estimated values.
  #
  # print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
  #
  # Plot the true and estimated CDFs.
  #      
  # curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
  # curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
  # points(x, pbeta(x, alpha, beta))
}
parms

rbeta(1e4, parms[1], parms[2]) %>% median()
rbeta(1e4, parms[1], parms[2]) %>% quantile(.025)
rbeta(1e4, parms[1], parms[2]) %>% quantile(.975)

rbeta(1e3, parms[1], parms[2]) %>% mean()

@@

>>>>

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near 0.2.", fig.width=9}

@@

@@

## Compare Priors

source(here("R/05-state-analysis.R"))

beta_shape = get_shape(ctis_smoothed,
                             option="beta",
                       quiet=TRUE)

tibble(x = seq(0, 1, length = 10^(4)),
       original = with(prior_params, beta_density(x,
                                     mean = beta_mean,
                                     sd = beta_sd,
                                     bounds = beta_bounds)),
       ctis= dbeta(x, shape1=beta_shape[1], 
                   shape2=beta_shape[2])) %>%
  pivot_longer(c(original,ctis))%>%
   ggplot(aes(x=x, y = value, fill=name)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0, ymax=value,fill=name),
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\beta$"),
       fill ='')

s_untested_shape = get_shape(ctis_smoothed,
                             option="s_untested",
                       quiet=TRUE)

tibble(x = seq(0, 1, length = 10^(4)),
       original = with(prior_params, beta_density(x,
                                     mean = s_untested_mean,
                                     sd = s_untested_sd,
                                     bounds = s_untested_bounds)),
       ctis= dbeta(x, shape1=s_untested_shape[1], 
                   shape2=s_untested_shape[2])) %>%
  pivot_longer(c(original,ctis))%>%
   ggplot(aes(x=x, y = value, fill=name)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0, ymax=value,fill=name),
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $Pr(S_1|untested)$"),
       fill ='')
